### 2017-03-14 CO750-ApproximationAlgorithmsTopics/assignments/assignment_02.tex
q5

%Q5
\section{}
\paragraph{}
Academic Honesty Disclaimer: In this problem I rely heavily on her paper Dr. Sanita gave us on Learn: Steiner Tree Approximation via Iterative Randomized Rounding.
\paragraph{}
Given a metric graph $G=(V,E)$ with edge costs $c:E\rightarrow \R_+$ and a set of terminals $R\subseteq V$, the $k$-restricted Steiner Tree problem asks for a tree of minimum cost spanning $R$ whose full components each contain at most $k$ terminals. From class the following algorithm was shown to be a $\ln(4)$-approximation algorithm for this problem.
\paragraph{}
Let $\cC$ denote the set of all directed components which can be obtained in the following way. Let $R' \subseteq R$ such that $|R'| \leq k$. and let $r' \in R'$. Let $C$ be the minimum cost Steiner tree on $R'$, and direct all edges of $C$ towards $r'$. Then $r'$ denotes the unique sink terminal in this case. A component $C \in \cC$ ``crosses" $U\subseteq R$ if $C$ has at least one source in $U$ and its unique sink in $R\backslash U$. The set $\delta^+(U)$ will denote the set of $C \in \cC$ which cross $U$. Fix an arbitrary root terminal $r \in R$. The directed cut relaxation $(k-DCR)$ is
\begin{align*}
\min \sum_{C\in \cC} c(C)x_C& \\
\text{s.t.} \sum_{C\in\delta^+(U)} x_C &\geq 1 &\forall U\subseteq R\backslash \{r\}, U\neq \emptyset \\
x&\geq 0.
\end{align*}
As argued in class this is a relaxation of the $k$-restricted Steiner tree problem. The iterative randomized rounding algorithm, $\cA$, for this problem is:
\begin{enumerate}
\item For $t =1,2,\dots$
\begin{enumerate}
\item Solve $(k-DCR)$ on $G$ to obtain $x^t$.
\item Sample one component $C^t$ where $C^t=C$ with probability $x^t_C/\sum_{C\in\cC} x^t_C$. Contract $C^t$ into its sink in $G$.
\item If a single terminal remains, return $\bigcup_{i=1}^t C^i$.
\end{enumerate}
\end{enumerate}
\paragraph{}
Since we already know from class that $\cA$ is a $\ln(4)$-approximation algorithm for $k$-restricted Steiner Tree, it only remains to verify the new approximation factor of $\frac{5}{4}$ when we fix $k=4$. Let $S^*$ be an optimal $4$-restricted Steiner Tree with undirected full components $Z_1, \dots, Z_\ell$. We will start by studying the structure of each $Z_i$. The first observation, made in the paper, is that we may assume Steiner Nodes of $S^*$ have degree at least $3$. This follows since $G$ is a metric graph we may shortcut any Steiner Node of degree two by connecting its endpoints and removing it. By the triangle inequality this shortcutting operation does not increase the cost of the solution.
\paragraph{}
We now move to observations about each $Z_i$.
\begin{lemma}\label{lemma:dk}
If $Z_i$ is a $k$-component for any $k\leq 4$, then the degree of any Steiner Node in $Z_i$ is at most $k$.
\end{lemma}
\begin{proof}
Suppose for a contradiction that $Z_i$ has a Steiner Node $s$ of degree greater than $k$. Then $s$ has at least $k+1$ neighbours: $v_1, \dots, v_{k+1}$. Since $Z_i$ is a tree, there is a unique path from $s$ to each of the $k$ terminals of $Z_i$, $r_1, \dots, r_k$. Each path $P_{s,r_j}$ can use only one of $v_1,\dots, v_{k+1}$. If they use more, say first $v_i$ and then next $v_j$ then the subpath $r,v_i,\dots,v_j$ with the edge $v_js$ forms a cycle in $Z_i$ contradicting that $Z_i$ is a tree. But then one of $v_1, \dots, v_{k+1}$ is not used on any path from $s$ to any of $r_1, \dots, r_k$. The edge from $s$ to the unused $v_j$ can then be removed from $Z_i$ without disconnecting the terminals, contradicting optimality.
\end{proof}
\end{document}

>>>>

%Q5
Academic Honesty Disclaimer: In this problem I rely heavily on her paper Dr. Sanita gave us on Learn: Steiner Tree Approximation via Iterative Randomized Rounding.
Given a metric graph $G=(V,E)$ with edge costs $c:E\rightarrow \R_+$ and a set of terminals $R\subseteq V$, the $k$-restricted Steiner Tree problem asks for a tree of minimum cost spanning $R$ whose full components each contain at most $k$ terminals. From class the following algorithm was shown to be a $\ln(4)$-approximation algorithm for this problem.
Let $\cC$ denote the set of all directed components which can be obtained in the following way. Let $R' \subseteq R$ such that $|R'| \leq k$. and let $r' \in R'$. Let $C$ be the minimum cost Steiner tree on $R'$, and direct all edges of $C$ towards $r'$. Then $r'$ denotes the unique sink terminal in this case. A component $C \in \cC$ ``crosses" $U\subseteq R$ if $C$ has at least one source in $U$ and its unique sink in $R\backslash U$. The set $\delta^+(U)$ will denote the set of $C \in \cC$ which cross $U$. Fix an arbitrary root terminal $r \in R$. The directed cut relaxation $(k-DCR)$ is
\min \sum_{C\in \cC} c(C)x_C& \\
\text{s.t.} \sum_{C\in\delta^+(U)} x_C &\geq 1 &\forall U\subseteq R\backslash \{r\}, U\neq \emptyset \\
As argued in class this is a relaxation of the $k$-restricted Steiner tree problem. The iterative randomized rounding algorithm, $\cA$, for this problem is:
\item For $t =1,2,\dots$
\item Solve $(k-DCR)$ on $G$ to obtain $x^t$.
\item Sample one component $C^t$ where $C^t=C$ with probability $x^t_C/\sum_{C\in\cC} x^t_C$. Contract $C^t$ into its sink in $G$.
\item If a single terminal remains, return $\bigcup_{i=1}^t C^i$.
Since we already know from class that $\cA$ is a $\ln(4)$-approximation algorithm for $k$-restricted Steiner Tree, it only remains to verify the new approximation factor of $\frac{5}{4}$ when we fix $k=4$. Let $S^*$ be an optimal $4$-restricted Steiner Tree with undirected full components $Z_1, \dots, Z_\ell$. We will start by studying the structure of each $Z_i$. The first observation, made in the paper, is that we may assume Steiner Nodes of $S^*$ have degree at least $3$. This follows since $G$ is a metric graph we may shortcut any Steiner Node of degree two by connecting its endpoints and removing it. By the triangle inequality this shortcutting operation does not increase the cost of the solution.
We now move to observations about each $Z_i$.
\begin{lemma}\label{lemma:dk}
If $Z_i$ is a $k$-restricted component for any $k$, then the degree of any Steiner Node in $Z_i$ is at most $k$.
Suppose for a contradiction that $Z_i$ has a Steiner Node $s$ of degree greater than $k$. Then $s$ has at least $k+1$ neighbours: $v_1, \dots, v_{k+1}$. Since $Z_i$ is a tree, there is a unique path from $s$ to each of the $k$ terminals of $Z_i$, $r_1, \dots, r_k$. Each path $P_{s,r_j}$ can use only one of $v_1,\dots, v_{k+1}$. If they use more, say first $v_i$ and then next $v_j$ then the subpath $r,v_i,\dots,v_j$ with the edge $v_js$ forms a cycle in $Z_i$ contradicting that $Z_i$ is a tree. But then one of $v_1, \dots, v_{k+1}$ is not used on any path from $s$ to any of $r_1, \dots, r_k$. The edge from $s$ to the unused $v_j$ can then be removed from $Z_i$ without disconnecting the terminals, contradicting optimality.
We now consider the classes of graphs these components appear as. It is an easy observation that $1$-restricted components only appear in the trivial case where $|R| = 1$. It is also easy to see that $2$-restricted components arise as a single edge connecting two terminals since Steiner Nodes have degree at least three. Further we see that $k$-restricted components with exactly one Steiner Node are star graphs centred at the Steiner Node. By lemma \ref{lemma:dk} $3$-restricted components have a Steiner Nodes with degree exactly three, and thus they can only have one Steiner Node. This implies that $3$-restricted components are star graphs. Things get slightly more interesting with $4$-restricted components.
\begin{lemma}\label{lemma:four}
A $4$-restricted component $Z_i$ is either a star graph or consists of exactly two adjacent degree three Steiner Nodes each adjacent to precisely two distinct terminals.
Suppose $Z_i$ is not a star graph. Then $Z_i$ has more than one Steiner Node. We claim that $Z_i$ has at most two adjacent Steiner Nodes. Suppose for a contradiction $Z_i$ has at least three adjacent Steiner Nodes. Say $s_1,s_2,s_3$ is a path of Steiner Nodes in $Z_i$. If the removal of edge $s_1s_2$ separates one terminal in the component containing $s_1$ then $s_1$ is adjacent to a redundant edge since its degree is at least three. So the edge removal separates at least two terminals in the component containing $s_1$. Similarly, by removing $s_2s_3$ we see that there are two terminals in the component containing $s_3$. But then, since $s_2$ of degree at least three, any third component it connects to (that is, not containing $s_1$ or $s_3$) is redundant to connect the terminals. This violates optimality. Hence $Z_i$ contains exactly two Steiner Nodes, and we see by the above argument that the removal of their adjacent edge separates two terminals into each connected component.
We are prepared to analyze the algorithm. Following the notation in the paper we let $D(e) = \max\{t : e \in S^t\}$ be the iteration at which $e$ is deleted. We construct the witness tree via the following procedure:
\item Start with $\tilde{B} = \emptyset$
\item For each $i = 1,\dots, \ell$
	\item If $Z_i$ is a $2$-restricted component add its edge to $\tilde{B}$.
	\item If $Z_i$ is a $3$-restricted component or $4$-restricted star graph component add all the edges of $Z_i$ but one chosen uniformly at random.
	\item If $Z_i$ is a $4$-restricted component with two adjacent Steiner Nodes $s_1,s_2$ and terminals $r_1, r_2$ connected to $s_1$ and terminals $r_3,r_4$ connected to $s_2$ then:
		 \item add edge $s_1s_2$ to $\tilde{B}$
		 \item add one edge of $r_1s_1$, $r_2s_1$ to $\tilde{B}$ chosen uniformly at random
		 \item add one edge of $r_3s_2$, $r_4s_2$ to $\tilde{B}$ chosen uniformly at random
\item Return Witness Tree $W= \{\{u,v\} in {R\choose 2} : |P_uv \cap \tilde{B}| = 1\}$.
By lemma \ref{lemma:four} and the preceding discussion, this procedure covers all possible forms the components $Z_i$ can take. Further one can verify that $W$ is a terminal spanning tree. 
For any $e \in E(S^*)$ we define $W(e) = \{ uv \in E(W): e \in P_{uv}\}$. When the edges of $W(e)$ are $\textit{marked}$ (see paper) the edge $e$ is deleted. We have the following lemma from the paper/class (rephrased in terms of $W(e)$ for simplicity):
\begin{lemma}\label{lemma:20}
Let $e \in E(S^*)$. Then the expected number of iterations until all the edges of $W(e)$ are $\textit{marked}$ satisfies $$E[X(W(e))] \leq H_{|W(e)|} \cdot M,$$
where $M$ is an upperbound on $|\cC|$ polynomial in $n$.
\begin{lemma}\label{lemma:de}
For any edge $e \in E(S^*)$, the expected number of iterations until $e$ is deleted is
$$E[D(e)] \leq \frac{5}{4} M.$$
Let $Z_i$ be the undirected full component containing $e$. We proceed by case analysis on the class of graph $Z_i$ belongs to. We have the following cases:
\item $Z_i$ is $2$-restricted,
\item $Z_i$ is $3$ or $4$-restricted star,
\item $Z_i$ is a $4$-restricted component with two degree three Steiner Nodes.
We begin with Case $1$. This case is trivial: with probability $1$, $e$ is the edge connecting the terminals. Thus by lemma \ref{lemma:20} $E[D(e)] = 1\cdot H_1 \cdot M = M \leq \frac{5}{4} M.$
Now we consider Case $2$. Suppose $Z_i$ is a $k$-restricted star where $k \in \{3,4\}$. Then with probability $\frac{1}{k}$ (when $e$ is the edge not in $\tilde{B}$), $|W(e)| = k-1$. With probability $1-\frac{1}{k}$), $|W(e)| = 1$. Thus by lemma \ref{lemma:20} we have
\begin{align*}E[D(e)] &\leq \frac{1}{k} H_{k-1} M + (1-\frac{1}{k})H_1 M
\\&=\begin{cases}
(\frac{1}{3}H_2 + \frac{2}{3})M, &\text{if $k = 3$} \\
(\frac{1}{4}H_3 + \frac{3}{4})M, &\text{if $k=4$}.
\end{cases} \\
&=\begin{cases}
\frac{7}{6}M, &\text{if $k = 3$} \\
(\frac{29}{24})M, &\text{if $k=4$}.
\end{cases} \\
&\leq \frac{5}{4} M.
Finally we consider Case $3$. By lemma \ref{lemma:four} we have the following structure.Let $s_1,s_2$ be the two adjacent Steiner Nodes of $Z_i$ and let $r_1,r_2$ be the terminals adjacent to $s_1$, and let $r_3,r_4$ be the terminals adjacent to $s_2$. Suppose without loss of generality that $s_1r_1, s_2r_4 \in \tilde{B}$. The the probability that $|W(e)| = 1$ is $\frac{3}{5}$ (such edges are $s_1r_1, s_2r_4$ and $s_1s_2$). The probability that $|W(e)| = 2$ is $\frac{2}{5}$ (such edges are $s_1r_2$ and $s_2r_3$). Thus by lemma \ref{lemma:20} we have
E[D(e)] &\leq \frac{3}{5}H_1M + \frac{2}{5}H_2M \\
&= (\frac{3}{5} + \frac{2}{10}) M \\
&= (\frac{8}{10})M \\
&= \frac{4}{5}M.
Therefore the expectation holds in each case as desired.
We now finish by computing the expected cost of the solution returned by $\cA$. Let $opt^t_f$ denote the optimal fractional cost to $(k-DCR)$ at iteration $t$. Let $opt$ denote the cost of $S^*$.
The algorithm $\cA$ is $\frac{5}{4}$-approximation algorithm for $4$-restricted Steiner Tree.
We observe that
E[\sum_{t\geq 1} c(C^t)] &\leq \sum_{t\geq 1} \frac{1}{M}E[opt^t_f] \\
&\leq \frac{1}{M} \sum_{t\geq 1} E[c(S^t)] \\
&=\frac{1}{M} \sum_{t\geq 1} E[D(e)]c(e) \\
&\leq \frac{1}{M} \sum_{t\geq 1} \frac{5}{4}M\cdot c(e) &\text{by lemma \ref{lemma:de}} \\
&=\frac{5}{4} \sum_{t \geq 1} c(e) \\
&= \frac{5}{4} opt.

%Q6
\section{}
\end{document}
